---
title: "4-fittsProcessing"
author: "Will Sheppard"
date: "2024-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries 1}

require(tidyverse)
require(rstatix)
```


```{r read data}
dataImport <- read_csv("../data/processed/fittsLong.csv")
head(dataImport)

pps <- read_csv("../data/finalParticipants.csv")
pps <- pps$x


```


```{r pos}


allData <- dataImport %>%
  filter(Participant.Private.ID %in% pps)%>%
  group_by(Participant.Private.ID)%>%
  mutate(id = as.factor(dplyr::cur_group_id()))%>%
  mutate(dist = sqrt(xR^2 + yR^2))%>%
  mutate(diffInd = log2((2*dist)/size))
# extract position from structure
allData$pos <- as.factor(str_extract(allData$structure, "(?<=pos)."))

```

First, lets look at some v basic distribution and box plots.
there seems to be some really extreme 
```{r initial plots}


ggplot(allData, aes(x = Reaction.Time, fill = as.factor(structure)))+
  geom_boxplot()

```

let's identify and plot the outlier values - 1.5 IQR above or below 3rd or first quartile
```{r is extreme}

extremePoints <- allData %>%
  group_by(id, Eye.Condition, structure) %>%
  mutate(ext = as.numeric(is_outlier(Reaction.Time)))


(sum(extremePoints$ext)/length(extremePoints$ext))*100

# moving through 4 layers of filtering to get rid of outliers. 
# first remove general outliers
# then get rid of outliers on the individual level (no task component)
# get rid of outliers on a task level (no indivisual component)
# get rid of any remaining outliers on a micro level
outliersRemoved <- allData %>% 
  # mutate(outlier = is_outlier(Reaction.Time))%>% # create logical markers for general outliers. this will get rid of the very extreme points
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0)%>%
  # group_by(id)%>% # this should catch any very strange behaviours at an individual level
  # mutate(outlier = is_outlier(Reaction.Time))%>% # 
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0)%>%
  # group_by(Eye.Condition, structure)%>% # this will catch strange behaviours at a task level
  # mutate(outlier = is_outlier(Reaction.Time))%>%
  # mutate(outlier = as.numeric(outlier))%>%
  # filter(outlier == 0) %>%
  group_by(id, Eye.Condition, structure)%>% # this will catch strange behaviours al the microlevel
  mutate(outlier = is_outlier(Reaction.Time))%>%
  mutate(outlier = as.numeric(outlier))%>%
  filter(outlier == 0) %>%
  ungroup()

percRem = (length(allData$Participant.Private.ID)-length(outliersRemoved$Participant.Private.ID))*100/length(allData$Participant.Private.ID)

# this removes 11% of data, which is a lot
outliersRemoved %>%
  ggplot(aes(x = as.factor(diffInd), y = Reaction.Time, color = Eye.Condition))+
    geom_boxplot()

```

let's remove the extreme data points: extra long movements suggest that they were not concentrating and extra short movemnts suggest that they haave guessed the posiitons
```{r tidy data set}

data <- outliersRemoved %>%
  select(id, Eye.Condition, diffInd, Reaction.Time, structure)%>%
  rename(visCond = Eye.Condition,
         rt = Reaction.Time
         )%>%
  mutate(visCondNum = case_when(
    visCond == 'Mono' ~ 1,
    TRUE ~ 0
  ))%>%
  mutate(diffInd = round(diffInd, digits = 2))

vs <- read.csv("./data/processed/visualSearchTimes.csv")

d <- merge(data, vs)%>%
  select(-X)

d$diffInd.s <- as.numeric(scale(d$diffInd))
d$visSearch.s <- as.numeric(scale(d$visSearch))

d$visCond <- as.factor(d$visCond)

str(d)
```


import libraries for mlm
```{r mlm libraries}
require(lme4)
require(lmerTest)
```


first we create a linear model with no random factors
no significant interaction
```{r linear mod}

# no interaction, remove
linMod <- lm(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s,
             data = d)
summary(linMod)

linModFinal <- linMod

```
Start with the maximal model and then remove factors starting with the least variance

glmer(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
                   (visCondNum*diffInd.s|id), # RANDOM
              family = inverse.gaussian(link = "identity"),
              data = data, 
              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000)))
              
it is reasonable to expect the slopes fo visual condition to vary between participants

no slope for diffInd as this has been shown to be a stable effect across individuals

the effect of search time on RT should be the same for all particpants as this is a direct relationship - longer ST = longer RT
```{r inv g}


# # bobyqa - Failed to converge
# # nloptwrap - failed to converge
# # Nelder_Mead - failed to converge
# invGMax <- glmer(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
#                    (visCondNum*diffInd.s|id), # RANDOM
#               family = inverse.gaussian(link = "identity"),
#               data = d, 
#               control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000)))
# summary(invGMax)

# bobyqa - Converged
# nloptwrap - didn't try
# Nelder_Mead - didn't try
invGTwo <- glmer(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
                   (visCondNum + diffInd.s|id), # RANDOM
              family = inverse.gaussian(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000)))
summary(invGTwo)


invGfinal <- invGTwo

residuals <- resid(invGfinal) 


qqnorm(residuals)

```


signularity when all possible factors are included. move through removing the slopes one by one from the smallest sd to the largest.
```{r gamma}

# # bobyqa - Singular
# # nloptwrap - doesn't converge
# # Nelder_Mead - doesn't converge
# gammaMax <- glmer(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
#                    (visCondNum*diffInd.s|id), # RANDOM
#               family = Gamma(link = "identity"),
#               data = d, 
#               control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000)))
# summary(gammaMax)
# isSingular(gammaMax)

# bobyqa - Converged
# nloptwrap - didn't try
# Nelder_Mead - didn't try
gammaTwo <- glmer(rt ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
                   (visCondNum + diffInd.s|id), # RANDOM
              family = Gamma(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000)))
summary(gammaTwo)

# bobyqa - failed to Converge
# nloptwrap - failed to Converge
# Nelder_Mead - failed to Converge
# gammaTwoNoST <- glmer(rt ~ visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
#                    (visCondNum + diffInd.s|id), # RANDOM
#               family = Gamma(link = "identity"),
#               data = d, 
#               control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000)))
# summary(gammaTwoNoST)

# bobyqa - failed to converge
# nloptwrap - is singular
# Nelder_Mead - failed to converge
# # (diffInd.s|id) has lowest varaince, remove
# gammaThreeNoST <- glmer(rt ~ visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
#                    (visCondNum|id) + (diffInd.s|id), # RANDOM
#               family = Gamma(link = "identity"),
#               data = d, 
#               control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000)))
# summary(gammaThreeNoST)

# bobyqa - Converged
# nloptwrap - did not try 
# Nelder_Mead - did not try
gammaFourNoST <- glmer(rt ~ visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
                   (visCondNum|id), # RANDOM
              family = Gamma(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000)))
summary(gammaFourNoST)

gammaFinal <- gammaTwo
gammaFinalNoST <- gammaFourNoST

d$rt.r <- round(d$rt, digits = -1)

gammaFinal.R <- glmer(rt.r ~ visSearch.s + visCondNum + diffInd.s + visCondNum*diffInd.s + #FIXED
                   (visCondNum + diffInd.s|id), # RANDOM
              family = Gamma(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "nloptwrap", optCtrl = list(maxfun = 10000000)))
summary(gammaFinal.R)


gammResiduals <- resid(gammaFinal) 


qqnorm(gammResiduals)
```

lets compare the three models
intercepts indicate performance when fixed factors = 0

intercept = mean visSEarch, mean diffInd and binocular vision
  1059.85

VIS SEARCH -  increase of 1 SD = +27.36 ms RT

VIS COND - 39.90*visCond - 3.70*visCond*diffInd
    monocular vision = 1, therefore, 39.90 - 3.70*diffInd
    
DIFF IND - 62.12*diffInd - 3.70*visCond*diffInd
  MONO - 62.12*diffInd - 3.70*diffInd
  BINO - 62.12*diffInd
```{r models}
require(modelsummary)
require(tibble)
require(flextable)

anova(invGfinal, gammaFinal)

modelsummary(list("Linear" = linModFinal, "Inverse Gaussian" = invGfinal, "Gamma" = gammaFinal),
             stars = T,
             title = "Clicking task model outpus",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             shape = term ~ model + statistic, #statistics in separate columns
             fmt = fmt_decimal(digits = 0, pdigits = 3),
             output = "fittsDistTable.html"
             )

modelsummary(list("Gamma" = gammaFinal, "Gamma No ST" = gammaFinalNoST),
             stars = T,
             title = "Clicking task model outpus",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             shape = term ~ model + statistic, #statistics in separate columns
             fmt = fmt_decimal(digits = 0, pdigits = 3),
             gof_omit = 'RMSE|IC|Obs',
             output = "fittsWithAndWithoutST.html"
             )

modelsummary(list("Gamma" = gammaFinal, "Rounded" = gammaFinal.R),
             stars = T,
             title = "Clicking task model outpus",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             shape = term ~ model + statistic, #statistics in separate columns
             fmt = fmt_decimal(digits = 0, pdigits = 3)
             )

```





```{r plots again}

fittsFig <- d %>%
  mutate(visCond = case_when(
    visCond == "Binoc" ~ "Binocular vision",
    visCond == "Mono" ~ "Monocular vision"
  ))%>%
  ggplot()+
  stat_summary(
    aes(y = rt, x = diffInd,
        group = visCond, color = visCond, shape = visCond),
    geom = "point",
    fun.y = "mean"
  )+
  geom_smooth(aes(y = rt, x = diffInd,
              group = visCond, color = visCond, linetype = visCond),
              method = "lm",
              se = F)+
  xlab("Index of difficulty (ID)")+
  ylab("Mean response time (ms)")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.background = element_blank(),
        legend.position = c(0.83, 0.28),
        text=element_text(family = "serif",
                          size=11))+
  scale_color_manual(values = c("grey70", "grey20"))+
  scale_shape_manual(values = c(16, 3))

show(fittsFig)
  
```



```{r save figure}

dpi = 800
w = 13
h = w*.75

ggsave('../figures/SUBMISSION/figure4.png',
       plot = fittsFig,
       dpi = dpi,
       width = w,
       height = h,
       units = "cm")

```





