---
title: "5-visSearchProcessing"
author: "Will Sheppard"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries 1}

require(tidyverse)
require(rstatix)
require(ggpubr)
```


```{r read data}
dataImport <- read_csv("../data/processed/vsLong.csv")
head(dataImport)

pps <- read_csv("../data/finalParticipants.csv")
pps <- pps$x

allData <- dataImport %>%
  filter(Participant.Private.ID %in% pps)%>%
  group_by(Participant.Private.ID)%>%
  mutate(id = as.factor(dplyr::cur_group_id()))
```

remove extra tasks
```{r remove tasks}

dataCorrect <- allData %>%
  filter(correct == 1)
```

quick first plots
```{r initial plots}

ggplot(dataCorrect, aes(x = Reaction.Time, fill = Eye.Condition))+
  geom_density(alpha = 0.4)+
  facet_grid(~VS_Puzzle)

ggplot(dataCorrect, aes(y = Reaction.Time, x = Eye.Condition, fill = VS_Puzzle))+
  geom_boxplot()

```

lets get rid of the extreme outliers

let's identify and plot the extreme values - 3IQR above or below 3rd or first quartile

There are no extreme data points. 
```{r is extreme}

extremePoints <- dataCorrect %>%
  group_by(id, Eye.Condition, VS_Puzzle)%>%
  mutate(ext = as.numeric(is_outlier(Reaction.Time)))

(sum(extremePoints$ext)/length(extremePoints$ext))*100

```

puzzle 3 = 0
4 = 1
1 = 2
2 = 3

```{r modelling data prep}

d <- dataCorrect%>%
  rename(visCond = Eye.Condition,
         rt = Reaction.Time
         )%>%
  mutate(visCondNum = case_when(
    visCond == 'Mono' ~ 1,
    TRUE ~ 0
  ))%>% ungroup() %>%
  mutate(puzzleNum = case_when(
    VS_Puzzle == "Image3.PNG" ~ 0,
    VS_Puzzle == "Image4.PNG" ~ 1,
    VS_Puzzle == "Image1.PNG" ~ 2,
    VS_Puzzle == "Image2.PNG" ~ 3
  ))%>%
  mutate(puzzleNum = as.factor(puzzleNum))

levels(d$puzzleNum)

d %>%
  shapiro_test(rt)

```

let extract a mean response time for each participant in each conditions
```{r mean extract, eval=FALSE}

means <- d %>%
  group_by(id, visCondNum)%>%
  summarise(visSearch = mean(rt))

write.csv(means, "./data/processed/visualSearchTimes.csv")
```

effect of visual condition approaching sig
```{r linear model}

linMod <- lm(rt ~ visCondNum,
             data = d)

summary(linMod)
```
import libraries for mlm
```{r mlm libraries}
require(lme4)
require(lmerTest)
```

RT data, so we will only use inverse gausian and gamma distributions.
fixed effects: eye condition
random int: id, puzzle
random slope: eye condition within id

invGmax  <- glmer(rt ~ visCondNum + (visCondNum|id) + (1|VS_Puzzle),
              family = inverse.gaussian(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
              )
              
  

```{r invG}

# failed to converge
# random slope has the lowest variance
invGmax  <- glmer(rt ~ visCondNum + (visCondNum|id) + (1|puzzleNum),
              family = inverse.gaussian(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
              )
summary(invGmax)

#invGtwo <- glmer(rt ~ visCondNum + (1|id) + (1|VS_Puzzle),
#              family = inverse.gaussian(link = "identity"),
#              data = d, 
#              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = #10000000))
#              )

#summary(invGtwo)


invGfinal <- invGmax

d$rt.r <- round(d$rt, digits = -1)

invGfinal.r <- glmer(rt.r ~ visCondNum + (1|id) + (1|puzzleNum),
              family = inverse.gaussian(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
              )
```


let's look at gamma
```{r gamma}

# doesn't converge
# puzzle has the lowest variance
# gammaMax <- glmer(rt ~ visCondNum + (visCondNum|id) + (1|VS_Puzzle),
#               family = Gamma(link = "identity"),
#               data = d, 
#               control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000000))
#               )
# summary(gammaMax)

gammaTwo <- glmer(rt ~ visCondNum + (visCondNum|id),
              family = Gamma(link = "identity"),
              data = d, 
              control = glmerControl(optimizer = "Nelder_Mead", optCtrl = list(maxfun = 10000000))
              )
summary(gammaTwo)

gammaFinal <- gammaTwo
```

lets compare the three models
intercepts indicate performance when fixed factors = 0
```{r models}
require(modelsummary)
require(tibble)
require(flextable)

anova(invGfinal, gammaFinal)

modelsummary(list("Linear" = linMod, "Inverse Gaussian" = invGfinal, "Gamma" = gammaFinal),
             stars = T,
             title = "Visual search model outpus",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             shape = term ~ model + statistic, #statistics in separate columns
             fmt = fmt_decimal(digits = 0, pdigits = 3),
             output = "vsDistTable.html"
             )

modelsummary(list("Inverse Gaussian" = invGfinal, "Rounded" = invGfinal.r),
             stars = T,
             title = "Visual search model outpus",
             estimate = "{estimate} [{conf.low}, {conf.high}]{stars}",
             statistic = "({std.error})",
             shape = term ~ model + statistic, #statistics in separate columns
             fmt = fmt_decimal(digits = 0, pdigits = 3)
             )

```




```{r paper plot}


vsFig <- d %>%
  group_by(visCond) %>%
  summarise(se = sd(rt, na.rm = TRUE)/sqrt(length(rt)),
            rt = mean(rt)) %>%
  ggplot(aes(x = visCond, y = rt, fill = visCond)) +
  geom_col(position = position_dodge(0.8), width = 0.7)+
  geom_errorbar(
    aes(ymin = rt-se, ymax = rt+se),
    width = 0.2, position = position_dodge(0.8)
    )+
  scale_fill_manual(values = c("grey80", "grey20"))+
  theme_bw()+
  scale_x_discrete(labels = c("Binocular vision", "Monocular vision"))+
  # set axis limits in coord_cartesian
 coord_cartesian(ylim = c(8000, 15000))+
  scale_y_continuous(breaks = seq(8000, 15000, by = 1000))+
  ylab("Mean search time (ms)")+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.position = "none",
        text = element_text(family = "serif",
                            size = 11))+
  geom_signif(stat="signif",position="identity",
                 comparisons=list(c("Binoc","Mono")),map_signif_level = TRUE,
              annotations = "***")

show(vsFig)
```

```{r save figure}

dpi = 800
w = 13
h = w*0.75

ggsave('../figures/SUBMISSION/figure3.png',
       plot = vsFig,
       dpi = dpi,
       width = w,
       height = h,
       units = "cm")

```