---
title: "8-perceptualLoad"
author: "Will Sheppard"
date: "2024-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create data frames
- this script aims to quantify the 'perceptual load' of each task to allow us to make the comparisons that we begin to introduce in the discussion. 

We speak about the Tsirlin et al. (2018) and the black et al. (2021) papers. 

We aim to quantify the relationship between the effect of visual condition and search time with regards to the number of targets presented to the participants at a given time.

We will achieve this by calculating the effect of visual conditions divided by the model intercept. In the case of the current paper this would be 1881.25/12509.25 = 0.1503887 or an increase of ~15.04%. 

However, it is not enough to simply compare this figure 

Black et al. - compare with TMT A task

```{r our data}

allDist <- 139

ourInt <- 12509.25
ourEffect <- 1881.25
ourIncPerc <- ourEffect*100/ourInt

ourEffectPerDist <- ourEffect/allDist
```